---
title: "All notes"
---

## Open files

Mac:

``` r
#| eval: false

Markdown02 <- "/Users/marcelorosales/Box Sync/Niigata Uni Box/Books-Notes/Notes/Markdown02.docx"  

Research_map <- "/Users/marcelorosales/Box Sync/Niigata Uni Box/Books-Notes/Notes/Research_map and guide.docx"  

Markdown01 <- "/Users/marcelorosales/Box Sync/Niigata Uni Box/Books-Notes/Notes/Markdown01.docx"

Notes <- "/Users/marcelorosales/Box Sync/Niigata Uni Box/Books-Notes/Notes/"

#open in mac OS
system(paste("open", shQuote(Markdown02)))
system(paste("open", shQuote(Research_map)))
system(paste("open", shQuote(Notes)))
```

Windows:

``` r
#| eval: false

file_path <- "/Users/marcelorosales/Box Sync/Niigata Uni Box/Books-Notes/Notes/Markdown01.docx"
shell(paste("start", shQuote(file_path)))
# or 
system(paste("start", shQuote(file_path)))
```

(See Markdown02.docx \> Word and Rstudio for detail info on the code)

## Files related to markdown. :

Spotlight search: Markdow

**Packages installed in R?**.

[installed_previously.csv](/Users/marcelorosales/Desktop/temp2022027/Box%20Sync/Documents/R/Rmarkdown/installed_previously.csv)\
installed_Rpackages20201004.csv.\
installed_Rpackages20201004.csv.\
installed_Rpackages20210108.csv.\
installed_Rpackages20210108.csv.\
oldRpks.csv.\
oldRpks.csv.

**Notes**\
docx2md_test.docx\
Does Decalcification speed.docx\
GeneExpression analysis.docx\
R Graph Notes.docx\
R Graph Notes.docx\
R Markdown notes.docx\
R Markdown notes.docx\
rmarkdown-brew.docx\
rmarkdown-brew.docx\
Untitled2.docx\
Untitled03.docx\
Untitled04.docx\
reference-keys.txt\
mkdocs.yml\
deforges.pdf\
Introductory R example chapters.pdf\
Introductory R example chapters.pdf rmarkdown-2.0.pdf\
rmarkdown-2.0.pdf\
rmarkdown-cheatsheet.pdf\
rmarkdown-cheatsheet.pdf\
teachR.pdf\
teachR.pdf\
Untitled.pdf

[top](#)

``` r
#| eval: false

Markdown02 <- "/Users/marcelorosales/Box Sync/Niigata Uni Box/Books-Notes/Notes/Markdown02.docx"  

Research_map <- "/Users/marcelorosales/Box Sync/Niigata Uni Box/Books-Notes/Notes/Research_map and guide.docx"  

Markdown01 <- "/Users/marcelorosales/Box Sync/Niigata Uni Box/Books-Notes/Notes/Markdown01.docx"

Notes <- "/Users/marcelorosales/Box Sync/Niigata Uni Box/Books-Notes/Notes/"

#open in mac OS
system(paste("open", shQuote(Markdown02)))
system(paste("open", shQuote(Research_map)))
system(paste("open", shQuote(Notes)))
```

|       | VS code |     |     | Rstudio |     |            |     |            |
|-------|---------|-----|-----|---------|-----|------------|-----|------------|
| code  | md      | Rmd | qmd | md      | Rmd | Rmd visual | qmd | qmd visual |
| r     | o       | o   | o   |         |     | o          |     | o          |
| .r    |         |     |     |         |     |            |     | o          |
| {r}   |         |     | o   | o       | o   |            | o   |            |
| {.r}  |         |     | o   | o       |     | o          |     | o          |
| rev r |         |     |     |         |     | o          |     | o          |

https://tableconvert.com/

| **VS code** | \*\*\*\* | \*\*\*\* | **Rstudio** | \*\*\*\* | \*\*\*\* | \*\*\*\*   | \*\*\*\* |
|-------------|----------|----------|-------------|----------|----------|------------|----------|
| **code**    | md       | Rmd      | qmd         | md       | Rmd      | Rmd visual | qmd      |
| **r**       | o        | o        | o           |          |          | o          |          |
| **.r**      |          |          |             |          |          |            |          |
| **{r}**     |          |          | o           | o        | o        |            | o        |
| **{.r}**    |          |          | o           | o        |          | o          |          |
| **rev r**   |          |          |             |          |          | o          |          |

```         
VS code         Rstudio             
```

code md Rmd qmd md Rmd Rmd visual qmd qmd visual r o o o o o .r o {r} o o o o\
{.r} o o o o rev r o o

```         
VS code VS code VS code Rstudio             
```

code md Rmd qmd md Rmd Rmd visual qmd qmd visual r o o o o o .r o {r} o o o o\
{.r} o o o o rev r o o

```         
VS code         Rstudio             
```

code md Rmd qmd md Rmd Rmd visual qmd qmd visual r o o o o o .r o {r} o o o o\
{.r} o o o o rev r o o

```{r}
#| eval: false
library(readxl)
highlight_long <- read_excel("~/Box Sync/Niigata Uni Box/Books-Notes/Notes/quarto_All.xlsx", 
    sheet = "Sheet2")
View(quarto_All)

library(tidyr)
highlight_wide <- spread(highlight_long, Soft, file, chunk, color)

```

```{r}
library(tidyverse)

# Create a data frame with the long-form table
df <- data.frame(
  Soft = c("VS code", "VS code", "VS code", "VS code", "VS code", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio", "Rstudio"),
  file = c("md", "md", "md", "md", "md", "md", "md", "md", "md", "md", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd", "Rmd"),
  chunk = c("r", ".r", "{r}", "{.r}", "rev r", "r", ".r", "{r}", "{.r}", "rev r", "r", ".r", "{r}", "{.r}", "rev r", "r", ".r", "{r}", "{.r}", "rev r", "r", ".r", "{r}", "{.r}", "rev r", "r", ".r", "{r}", "{.r}", "rev r"),
  color = c(1, 0, 0, 0, 0, 1, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 1, 0, 1, 0, 1, 1, 1)
)

# Convert the table to wide form
df_wide <- df %>%
  pivot_wider(names_from = chunk, values_from = color)

# Print the resulting wide-form table
print(df_wide)
```
