---
title: "R and Rstudio Install, setup and Package Management"
description: "How to install and upgrade R and RStudio"
author: "Marcelo Rosales"
date: "2023/03/24"
categories:
  - R
  - RStudio
  - Install
  - Rpackage
format: 
  html: 
    code-fold: false
    code-copy: true
    code-overflow: wrap
---

## Installing/Upgarading R and RStudio

References Pages to visit first:\
[Painless way to install a new version of R?](https://stackoverflow.com/questions/1401904/painless-way-to-install-a-new-version-of-r)\
[How to migrate Rstudio files and installed packages ( by version to a new computer )](https://stackoverflow.com/questions/67660414/how-to-migrate-rstudio-files-and-installed-packages-by-version-to-a-new-comput)\
[An efficient way to install and load R packages](https://statsandr.com/blog/an-efficient-way-to-install-and-load-r-packages/#most-efficient-way)\
[Loading packages: the difference between R's library() and require() functions](https://timfarewell.co.uk/loading-packages-the-difference-between-rs-library-and-require-functions/)\
[Bioconductor](https://www.bioconductor.org/install/)\
[Install/Update R and R packages](https://yiweiniu.github.io/blog/2019/07/Install-Update-R-and-R-packages/)\
[restore all R packages after installing a new version of R?](https://stackoverflow.com/questions/65817504/restore-all-r-packages-after-installing-a-new-version-of-r)\
[How to Upgrade R Without Losing Your Packages](https://www.r-bloggers.com/2015/06/how-to-upgrade-r-without-losing-your-packages/)\
[Upgrade R Without Losing Your Packages](https://www.r-bloggers.com/2017/05/upgrade-r-without-losing-your-packages/)\
[Update and migrate R](https://gist.github.com/arcaravaggi/20acc42e3e245cc268a95269de024b2d)\
[Profiles package](https://github.com/jmgirard/profiles)\
[Rprofile.site file example for Startup customization](https://www.statmethods.net/interface/customizing.html)\
[Managing R with .Rprofile, .Renviron, Rprofile.site, Renviron.site, rsession.conf, and repos.conf](https://support.posit.co/hc/en-us/articles/360047157094-Managing-R-with-Rprofile-Renviron-Rprofile-site-Renviron-site-rsession-conf-and-repos-conf)

Trick for doing the upgrade, from installing the software to copying all the settings/packages over?First you install the new version, then run this in the old version:

### Run in the old version of R

```{r}
#| eval: false  
setwd("C:/Temp/")
packages <- installed.packages()[,"Package"]
save(packages, file="Rpackages")
```

Followed by this in the new version:

### Run in the new version

```{r}
#| eval: false  
setwd("C:/Temp/")
load("Rpackages")
for (p in setdiff(packages, installed.packages()[,"Package"]))
install.packages(p)
```

### For windows

Using Installr

```{r}
#| eval: false  

#From https://stackoverflow.com/questions/1401904/painless-way-to-install-a-new-version-of-r

# Run in the old version of R (or via RStudio)
setwd("C:/Temp/") 
packages <- installed.packages()[,"Package"] 
save(packages, file="Rpackages") 

# INSTALL NEW R VERSION
if(!require(installr)) { install.packages("installr"); require(installr)} #load / install+load installr
# See here for more on installr: https://www.r-statistics.com/2013/03/updating-r-from-r-on-windows-using-the-installr-package/  

# step by step functions:
check.for.updates.R() # tells you if there is a new version of R or not.
install.R() # download and run the latest R installer... This install.R seems to be a windows only command. 

# Install library - run in the new version of R. This calls package names and installs them from repos, thus all packages should be correct to the most recent version
setwd("C:/Temp/") 
load("Rpackages") 
for (p in setdiff(packages, installed.packages()[,"Package"])) 
install.packages(p) 

# Installr includes a package migration tool but this simply copies packages, it does not update them
copy.packages.between.libraries() # copy your packages to the newest R installation from the one version before it (if ask=T, it will ask you between which two versions to perform the copying)
```

## Upgrading R on Windows and Mac

https://www.r-statistics.com/tag/installr/ 

### Windows

If you are using **Windows** you can easily upgrade to the latest version of R using the installr package. Simply run the following code in Rgui:

```{r}
#| eval: false  

install.packages("installr") # install 
setInternet2(TRUE) # only for R versions older than 3.3.0
installr::updateR() # updating R.
# If you wish it to go faster, run: installr::updateR(T)
```

Running "updateR()" will detect if there is a new R version available, and if so it will download+install it (etc.). There is also a step by step tutorial (with screenshots) on how to upgrade R on Windows, using the installr package. If you only see the option to upgrade to an older version of R, then change your mirror or try again in a few hours (it usually take around 24 hours for all CRAN mirrors to get the latest version of R).

### Mac

If you are using **Mac** you can easily upgrade to the latest version of R using Andrea Cirillo's updateR package. The package is not on CRAN, so you'll need to run the following code in Rgui:

```{r}
#| eval: false  

install.packages("devtools")
devtools::install_github("AndreaCirilloAC/updateR")
updateR(admin_password = "PASSWORD") # Where "PASSWORD" stands for your system password
```

MBP15 The downloaded source packages are in **"/private/var/folders/xc/rxn5vywj3lvd_qt78ss92cc80000gn/T/Rtmp8TJvA5/downloaded_packages"** MBP20 The downloaded binary packages are in **/var/folders/1v/nbx_5tgn553dq8s639lyln5c0000gn/T//Rtmp8AkWwy/downloaded_packages**

Option 1

[Painless way to install a new version of R?](https://stackoverflow.com/questions/1401904/painless-way-to-install-a-new-version-of-r)

Files are `.rda`

```{r}
#| eval: false  
#--run in the old version of R
setwd("C:/Temp/")
packages <- installed.packages()[,"Package"]
save(packages, file="Rpackages")

#--run in the new version
setwd("C:/Temp/")
load("Rpackages")
for (p in setdiff(packages, installed.packages()[,"Package"]))
install.packages(p)


#-----

# Just for completeness, you can save your packages in another directory on your computer.
install.packages("thepackage",lib="/path/to/directory/with/libraries")
#You can change the default .Library value using the function .libPaths too
.libPaths("/path/to/directory/with/libraries")

# -----

#Finally, You can also include a small code in my Rprofile.site allowing  to reinstall all packages when installing a new R version. You just have to list them up before you update to the new R versionbusing an .RData file containing an updated list with all packages.

library(utils)

## Check necessary packages
load("G:\Setinfo\R\packagelist.RData") # includes a vector "pkgs"
installed <- pkgs %in% installed.packages()[, 'Package']
if (length(pkgs[!installed]) >=1){
  install.packages(pkgs[!installed])
}
# I make the packagelist.RData by specifying .Last() in my Rprofile.site. This updates the package list if I installed some :

.Last <- function(){
  pkgs <- installed.packages()[,1]
  if (length(pkgs) > length(installed)){
    save(pkgs,file="G:\Setinfo\R\packagelist.RData")
  }
}

```

Option 2:\
[How to migrate Rstudio files and installed packages ( by version to a new computer )](https://stackoverflow.com/questions/67660414/how-to-migrate-rstudio-files-and-installed-packages-by-version-to-a-new-comput)\

Save as `.csv files`.\
Or `.xlsx files`.\

```{r}
#| eval: false  

getwd()
setwd()
# /Users/marcelorosales/Box Sync/Github/RStudio_shared/R_packages/file_name.csv  

installed <- as.data.frame(installed.packages())
write.csv(installed, 'installed_previously.csv') # Path/to/folder/installed_previously.csv 
installedPreviously <- read.csv('installed_previously.csv') # Path/to/folder/installed_previously.csv 
baseR <- as.data.frame(installed.packages())
toInstall <- setdiff(installedPreviously, baseR)

# the first thing to do would be to make sure that the set of currently installed packages is up-to-date. 

update.packages(checkBuilt=TRUE, ask=FALSE)  #check spelling of arguments
  new_pacs <- paste( setdiff( installedPreviously$Package, baseR$Package), 
                       collapse=",")
  install.packages( new_pacs, dependencies=TRUE)

# For excel files, requires: 
install.packages("writexl")
library("writexl")
write_xlsx(the_dataframe_name,"path/to/folder/file_name.xlsx")
  
```

## An efficient way to istall and load R packages

[An efficient way to install and load R packages](https://statsandr.com/blog/an-efficient-way-to-install-and-load-r-packages/#most-efficient-way)\

### Inefficient way to install and load R packages:\

```{r}
#| eval: false
# Installation of required packages
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("readxl")
install.packages("dplyr")

# Load packages
library("tidyverse")
library("ggplot2")
library("readxl")
library("dplyr")
```

### More efficient way

```{r}
#| eval: false

# Package names
packages <- c("ggplot2", "readxl", "dplyr", "tidyr", ...)

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
```

### Most efficient way

`{pacman} package`\
After this article was published, a reader informed me about the {pacman} package. After having read the documentation and try it out myself, I learned that the function p_load() from {pacman} checks to see if a package is installed, if not it attempts to install the package and then loads it. It can also be applied to several packages at once, all this in a very condensed way:

```{r}
#| eval: false
install.packages("pacman")

pacman::p_load(ggplot2, tidyr, dplyr)

## Make sure your current packages are up to date
update.packages()
## devtools is required
library(devtools)
install_github("trinker/pacman")

```

Find more about this package on [CRAN](https://github.com/trinker/pacman). [Introduction to the pacman Package in R (3 Examples)](https://statisticsglobe.com/pacman-r-package)\

`{librarian} package`\
Like `{pacman}`, the `shelf()` function from the `{librarian}` package automatically installs, updates, and loads R packages that are not yet installed in a single function. The function accepts packages from CRAN, GitHub, and Bioconductor (only if Bioconductor's Biobase package is installed). The function also accepts multiple package entries, provided as a comma-separated list of unquoted names (so no "" around package names).\

Last but not least, the `{librarian}` package allows to load packages automatically at the start of every R session (thanks to the `lib_startup()` function) and search for new packages on CRAN by keywords or regular expressions (thanks to the `browse_cran()` function).\

Here is an example of how to install missing packages and load them with the `shelf()` function:

```{r}
#| eval: false

# From CRAN:
install.packages("librarian")

librarian::shelf(ggplot2, DesiQuintans / desiderata, pander)
```

For CRAN packages, provide the package name as normal without "" and for GitHub packages, provide the username and package name separated by / (i.e., UserName/RepoName as shown for the desiderata package).

Find more about this package on CRAN.

## Location of the R packages


```{r}
#| eval: false
#| code-overflow: wrap
#| 
# _> In Terminal

# MBP15: The downloaded source packages are in:
open /private/var/folders/xc/rxn5vywj3lvd_qt78ss92cc80000gn/T/Rtmp8TJvA5/downloaded_packages 

# MBP20: The downloaded binary packages are in:
open  /var/folders/1v/nbx_5tgn553dq8s639lyln5c0000gn/T//Rtmp8AkWwy/downloaded_packages 

```


## Summary

#### Saved installed packages

```{r}
#| eval: false
#| 
# 20230324
# Update current packages.
update.packages(checkBuilt=TRUE, ask=FALSE)  #check spelling of arguments

# Upgrade packages of Bioconductor
# Install packages from a newer version of Bioconductor.
# https://www.bioconductor.org/install/ 

if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install(version = "3.16")

#--run in the old version of R
# MBP15: setwd("/Users/Marcelo-Rosales/Box Sync/Github/RStudio_shared/R_packages/")
RSpacks230324MBP15 <- installed.packages()[,"Package"]
save(RSpacks230324MBP15, file="/Users/Marcelo-Rosales/Box Sync/Github/RStudio_shared/R_packages/RSpacks230324MBP15")

#--run in the new version
# MBP20: setwd("/Users/marcelorosales/Box Sync/Github/RStudio_shared/R_packages/")

load("/Users/marcelorosales/Box Sync/Github/RStudio_shared/R_packages/RSpacks230324MBP15")
for (p in setdiff(RSpacks230324MBP15, installed.packages()[,"Package"]))
  install.packages(p)
```

#### Copy/Paste Rstudio global settings and preferences if necessary.

```{r}
#| eval: false
#| 
# Copy/Paste Rstudio global settings and preferences.. 

# In terminal _> 
open ~/.config/rstudio

# Search and Copy file to same place as packages files. 
rstudio-prefs.json
```

#### Brew cask upgrade R and Rstudio.

```{r}
#| eval: false
#| 
# IN _> Upgrade R and rstudio
brew update
brew upgrade --cask r
# password
#`brew cleanup` has not been run in the last 30 days, running now...

# Upgrade RStudio
brew upgrade --cask rstudio

# Copy paste the `rstudio-prefs.json` if needed, usually it still remains after upgrade. 
rstudio-prefs.json

```

#### Errors and trouble shooting. 

When opening R new version... an error message is displayed. 

```{r}
#| code-block-bg: false
#| code-block-border-left: "#31BAE9"
#| eval: false

During startup - Warning messages:
1: Setting LC_CTYPE failed, using "C" 
2: Setting LC_COLLATE failed, using "C" 
3: Setting LC_TIME failed, using "C" 
4: Setting LC_MESSAGES failed, using "C" 
5: Setting LC_MONETARY failed, using "C" 
[R.app GUI 1.79 (8198) x86_64-apple-darwin17.0]

WARNING: You're using a non-UTF8 locale, therefore only ASCII characters will work.
Please read R for Mac OS X FAQ (see Help) section 9 and adjust your system preferences accordingly.
```

A solution:
For english:
[Installing R on Mac - Warning messages: Setting LC_CTYPE failed, using "C"](https://stackoverflow.com/questions/9689104/installing-r-on-mac-warning-messages-setting-lc-ctype-failed-using-c)  

```{r}
#| eval: false
defaults write org.R-project.R force.LANG en_US.UTF-8  # 英語表示
defaults write org.R-project.R force.LANG ja_JP.UTF-8  # 日本語表示
```

[For Japanese:](https://maku77.github.io/r/basic/locale.html).  \




